Lending Club Exploratory Data Analysis
========================================================

Lending Club is an online lending service that allows borrowers to connect
with investors through a marketplace in a service known as peer-to-peer lending.
It's a form of lending that's known to often be cheaper than banks. The dataset
that will be explored in this exploratory data analysis investigation is one
that contains loan data for almost 900 000 loans handled by Lending Club between
2007 and 2015. It includes observations on over 70 variables including loan
description, interest rate, loan status, and many more. 

Through this analysis, we hope to gain a better understanding of:
1. What types of loans Lending Club are making?
2. What types of people take out these loans?
3. How does Lending Club create loans that are favourable for investors?


```{r, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(stringr)
library(gridExtra)
library(ggmap)
library(maps)
library(mapdata)
library(reshape2)

loan <- read.csv("loan.csv")

```

We notice right off the bat that this dataset is very large. We cannot possibly
examine every variable in this analysis. Returning to our primary aims, we can
narrow down the list of variables to examine.

It was also discovered that several key variables listed in the dataset
documentation are missing from the actual data. Most importantly, no FICO credit
information is available for any of the loans, taking away a key factor in 
the money lending process.

With this in mind, 15 variables have been selected from the larger dataset to
base this investigation upon. They have been placed into a separate datafram for
ease of use.

```{r}
working <- select(loan, addr_state, annual_inc, purpose, dti, grade, sub_grade,
                  issue_d, emp_title, home_ownership, loan_amnt, member_id,
                  term, int_rate, installment, delinq_2yrs, loan_status)
str(working)
```

# Univariate Plots and Analysis

We will start by examining the number of loans given over the duration of the
data recorded in the dataset.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Separate month and year from issue_d variable
working$issue_month <- factor(str_sub(working$issue_d, 1, 3))
working$issue_year <- strtoi(str_sub(working$issue_d, start = -4))

working$issue_month <- factor(working$issue_month,
                              levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                                         "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
                              ordered = TRUE)

# Group loans by year and month
date_sorted <- working %>%
  select(loan_amnt, issue_d, issue_month, issue_year) %>%
  group_by(issue_year, issue_month) %>%
  arrange(issue_year, issue_month) %>%
  summarise(count = n(), sum = sum(loan_amnt)) %>%
  mutate(date = paste(issue_month, issue_year, sep="-"))

ggplot(data = date_sorted, 
       aes(x = reorder(date, issue_year), y = count)) + 
  geom_bar(stat = "identity") +
  scale_x_discrete(breaks = c("Jul-2007", "Jul-2008", "Jul-2009", "Jul-2010",
                              "Jul-2011", "Jul-2012", "Jul-2013", "Jul-2014",
                              "Jul-2015"),
                   labels = c("2007", "2008", "2009", "2010", "2011", "2012",
                              "2013", "2014", "2015")) +
  labs(x = "Issue Date")

ggplot(data = date_sorted, 
       aes(x = reorder(date, issue_year), y = count)) + 
  geom_bar(stat = "identity") +
  scale_x_discrete(breaks = c("Jul-2007", "Jul-2008", "Jul-2009", "Jul-2010",
                              "Jul-2011", "Jul-2012", "Jul-2013", "Jul-2014",
                              "Jul-2015"),
                   labels = c("2007", "2008", "2009", "2010", "2011", "2012",
                              "2013", "2014", "2015")) +
  labs(x = "Issue Date") +
  scale_y_log10(breaks = c(100, 1000, 10000, 100000))

```

The number of loans given out increases steadily as the years pass. Looking at 
the second plot with a logarithmic y-axis, we can see this growth as an almost 
linear increase, indicating the number of loans made through Lending Club is 
increasing exponentially. We can also see that the growth is not constant. There
are many months were there are large spikes and other where there are large dips.
To see if there is any pattern in these spikes, we will look at the numbers 
behind the visualization.

```{r echo=FALSE, message=FALSE, warning=FALSE}

total_by_year <- by(date_sorted$count, date_sorted$issue_year, sum)
avg_per_month <- by(date_sorted$count, date_sorted$issue_month, mean)

print("Yearly totals:", quote = FALSE)
head(total_by_year, 9)

print("Monthly averages:", quote = FALSE)
head(avg_per_month, 12)

```

As expected, the number of loans made is increasing rapidly throughout the
years. In the 8 years between 2007 and 2015, the number of loans made through
Lending Club increased by almost 700 times.

Now the monthly average figures is where it gets interesting. General experience
would tell us that most people would probably take out loans most often around
the holiday season when many big purchases are made. However, there doesn't
appear to be any real, explainable trend in the monthly average data with the
two highest monthly averages occurring in October and July and the lowest in 
February and September.

***

Now that we know how many loans Lending Club are making, and when they are made,
we should learn about what *type* of loans they are facilitating. All loans made
are graded for stability and risk by lenders using factors like credit history, 
collateral, likelihood of repayment, etc. We can see the distribution of the
grades of the loans below.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = working, aes(grade)) + geom_histogram(stat = "count") +
  labs(y = "count", x = "grade")

ggplot(data = working, aes(sub_grade)) + geom_histogram(stat = "count") +
  labs(y = "count", x = "subgrade") 

```

As we can see, the distribution for both grade and subgrade is right-skewed
with most of the loans being of grade B or C. Looking at the more detailed 
subgrade classifications, we see the same general shape. Very high grade loans 
(A1, A2, A3) are not quite so common and the tail also tapers off quickly
with very few loans of grade F and G. This shows that most investors are more
likely to take on safer loans rather than make perhaps more money with more
risky options.

With this in mind, what is the risk like to investors? How many of Lending Clubs
loans are unsuccessful?

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = working, aes(loan_status)) + 
  geom_histogram(stat="count") +
  scale_x_discrete(limits = c("Current", "Fully Paid", "Charged Off",
                              "Late (31-120 days)", "Issued", "In Grace Period",
                              "Late (16-30 days)", "Default"),
                   labels = c("Current", "Fully\nPaid", "Charged\nOff",
                              "Late\n(31-120 days)", "Issued", "In Grace\nPeriod",
                              "Late\n(16-30 days)", "Default"))

ggplot(data = working, aes(loan_status)) + 
  geom_histogram(stat="count") +
  scale_x_discrete(limits = c("Current", "Fully Paid", "Charged Off",
                              "Late (31-120 days)", "Issued", "In Grace Period",
                              "Late (16-30 days)", "Default"),
                   labels = c("Current", "Fully\nPaid", "Charged\nOff",
                              "Late\n(31-120 days)", "Issued", "In Grace\nPeriod",
                              "Late\n(16-30 days)", "Default")) +
  scale_y_log10()

summary(working$loan_status)

```

The distribution shows that the vast majority (70%) of Lending Club loans are
currently in progress and up to date on all payments. The numbers then drop off
steeply leading into a very long tail that has been adjusted using a log(10) 
y-axis in the second plot. 

It would appear that most loans done through Lending Club are fairly safe for
investors. Out of 285 000 completed loans, 80% of them have been fully paid off
with only 0.4% of loans defaulting (120+ days past due) and around 16% charged
off (150+ days past due and with little expectation of repayment).

Additionally, to further understand the investment potential on Lending Club
loans, we should look at the potential returns for the investors. Lending Club
assigns an interest rate to each loan based on loan grade and borrower credit
history.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = working, aes(int_rate)) + geom_histogram() +
  geom_vline(xintercept = median(working$int_rate), color = "red") +
  geom_vline(xintercept = mean(working$int_rate), color = "blue") +
  labs(x = "interest rate (%)")

summary(working$int_rate)

```

The red line represents the median value and the blue line the mean. With the
mean, median, and mode all centered around the 13% mark, the distribution looks 
almost normal between 5% and 20% with a sharp fall-off after 20%. In fact, only 
7% of loans had interest rates above 20%.

***

Now let's look at the actual amount of money loaned out through Lending Club
services.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = working, aes(loan_amnt)) + geom_histogram() +
  labs(x = "loan amount ($)") + 
  geom_vline(xintercept = median(working$loan_amnt), color = "red") +
  geom_vline(xintercept = mean(working$loan_amnt), color = "blue")

summary(working$loan_amnt)

```

We notice from this plot the distinct peaks at "nice" amounts like \$10 000, 
\$15 000, and \$20 000, which is to be expected. The red line in the above plot 
indicates the median of \$13 000 and the blue line the mean of \$14 760. This 
indicates an overall right-skewed distribution with a larger number of loans
smaller than the mean rather than larger. We can see the distribution tapering 
into the higher values with, however, another peak at what appears to be a hard 
cap on maximum loans allowed to be given out of \$35 000.

With these loan amounts in mind, we will now examine the annual income of loan
applicants.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = working, aes(annual_inc)) + geom_histogram(binwidth=5000) +
  geom_vline(xintercept = median(working$annual_inc, na.rm=TRUE), color = "red") +
  geom_vline(xintercept = mean(working$annual_inc, na.rm=TRUE), color = "blue") +
  labs(x = "annual income ($)") +
  coord_cartesian(
    xlim = c(0, quantile(working$annual_inc, probs = 0.99, na.rm = TRUE))) 

ggplot(data = working, aes(annual_inc)) + geom_histogram(bins=60) +
  scale_x_log10(breaks = c(10000, 100000), 
                labels = c("$10 000", "$100 000")) +
  labs(x = "annual income") +
  coord_cartesian(
    xlim = c(quantile(working$annual_inc, probs = 0.01, na.rm = TRUE), 
             quantile(working$annual_inc, probs = 0.99, na.rm = TRUE)))

summary(working$annual_inc)

```

Once again, the red line represents the median and the blue, the mean. 
As we can see from these plots, the distribution for the annual income is right-
skewed with a long tail and extreme outliers (a maximum value of \$9 500 000).
Transformed with a log(10) x-axis, the distribution appears to be somewhat
normal. This means that there appear to be individuals with very high annual
incomes (into the high 6-figures and beyond) that are applying for relatively
small loans (up to a maximum of \$35 000). We should now look at who exactly the
people taking out these loans are.

The variable that would best tell us this is *emp_title*, which 
stores the self-declared job title of the applicant at the time of the loan 
application. However, according to the documentation, this variable had its 
purpose changed part-way through the data in the dataset. For entries up until 
09/23/2013, this variable didn't record the job title but the company name of 
the applicant's employer. Therefore, for the sake of consistency for this plot, 
only entries after the change are considered. This reduces the size of our
sample by more than 75%.

```{r message=FALSE, warning=FALSE}

# Select only data from after September 2013
after_sep_2013 <- subset(working, !is.na(emp_title)) %>%
  select(emp_title, issue_month, issue_year) %>%
  filter((issue_year > 2013) | (issue_year == 2013 & 
                                (issue_month %in% c("Oct", "Nov", "Dec")))) %>%
  group_by(emp_title) %>%
  summarise(n = n()) %>%
  arrange(desc(n))

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
job_titles <- after_sep_2013[2:11,]
job_titles$emp_title <- reorder(job_titles$emp_title, -job_titles$n)

ggplot(data = job_titles, aes(x = emp_title, y = n)) + 
  geom_bar(stat="identity") +
  labs(x = "job title", y = "count") +
  scale_x_discrete(labels =
    c("Teacher", "Manager", "Registered\nNurse", "Owner", "RN", 
      "Supervisor", "Sales", "Project\nManager", "Driver", "Office\nManager"))

```

From this chart it appears that teachers and managers are by far the most common
applicants for loans from Lending Club. There doesn't seem to be any real 
correlation or similarity between these 10 professions. There seems to be more 
management and leadership-type people in these categories but the remaining 
professions don't seem to share much, if anything, in common. There isn't much 
to indicate what types of people are the most likely to take out loans. This may 
be due to the earlier discussed discrepancy and smaller sample size, but we 
should look at how the applicants used their loans to see if there isn't another
factor at play.

```{r echo=FALSE, message=FALSE, warning=FALSE}

purpose <- subset(working, !is.na(purpose)) %>%
  select(purpose) %>%
  group_by(purpose) %>%
  summarise(n = n()) %>%
  arrange(desc(n))

purpose$purpose <- reorder(purpose$purpose, -purpose$n)

ggplot(data = purpose[1:10,], aes(x = purpose, y = n)) + 
  geom_bar(stat="identity") +
  scale_x_discrete(labels = 
    c("debt\nconsolidation", "credit\ncard", "home\nimprovement", "other",
      "major\npurchase", "small\nbusiness", "car", "medical", "vacation", "house"))

```

It would seem that the vast majority of loan applicants applied for the purpose 
of debt consolidation, with a smaller, but still significant number for credit 
card debts. Nothing here is really out of the ordinary or unexpected. The three
most common purposes are also the same ones advertised right on the front page
of Lending Club's website and are probably the most common sources of debt in
the US.

# Bivariate Plots and Analysis

While looking at the variables in the dataset individually has given us
important insights into the data, we can't really go about understanding what
the data can really tell us without looking at variables in relation to each
other.

We'll start by examining the geography of borrowers. (It's expected that there
won't be much learned about the data through this, but I wanted a chance to
explore the mapping packages offered in R.)

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Map code adapted from http://eriqande.github.io/rep-res-web/lectures/making-maps-with-R.html
# Population data taken from https://raw.githubusercontent.com/BuzzFeedNews/2015-11-refugees-in-the-united-states/master/data/census-state-populations.csv

states <- map_data("state")
state_pop <- read.csv("census-state-populations.csv")
colnames(state_pop) <- c("addr_state", "population")

state_data <- working %>%
  select(addr_state, loan_amnt) %>%
  group_by(addr_state) %>%
  summarise(count = n(), total = sum(loan_amnt))

state_data$addr_state <- state.name[match(state_data$addr_state, state.abb)]
state_data <- merge(state_data, state_pop, by = "addr_state")
state_data$addr_state <- tolower(state_data$addr_state)
colnames(state_data) <- c("region", "count", "total", "population")
state_data <- inner_join(states, state_data, by = "region")

# Remove axes but leave legend
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank())

ggplot(data = state_data) +
  geom_polygon(aes(x = long, y = lat, fill = total, group = group), 
               color = "black") +
  coord_fixed(1.3) +
  theme_bw() +
  ditch_the_axes +
  scale_fill_gradient(trans = "sqrt")

ggplot(data = state_data) +
  geom_polygon(aes(x = long, y = lat, fill = total / population, group = group), 
               color = "black") +
  coord_fixed(1.3) +
  theme_bw() +
  ditch_the_axes

```

It seems from this map that the highest loan amounts come from the states of
California, New York, and Texas. This is to be expected as these are also the
three most populous US states. After adjusting the map to show loan amounts
per person, there appears to be no real relationship between states and Lending
Club usage.

***

As a public company with shareholders to please, one of Lending Club's primary 
objectives is to be profitable. Part of this revolves around creating the most
favourable conditions for their investors to make returns. In this part of the 
investigation, we would like to determine how Lending Club assesses and sets 
the conditions of a loan (interest rate, grade, term, approved loan
amount, etc.).

We start by noting that Lending Club loans are offered in two different lengths
of terms, 36 months or 60 months. Let's examine these two types of loans
and how they differ.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = working, aes(x = term, y = int_rate)) +
  geom_boxplot() +
  labs(y = "interest rate (%)")

ggplot(data = working, aes(x = term, y = loan_amnt)) +
  geom_boxplot() +
  labs(y = "loan amount ($)")

# Function to determine percentage breakdowns of var1 within each
# value of var2

percent_breakdown <- function(data, var1, var2) {
  
  sum_by <- data %>%
    select_(var1, var2) %>%
    group_by_(var2, var1) %>%
    summarise(count = n())
  
  count_by <- by(sum_by$count, sum_by[[var2]], sum)

  count_by<- sapply(count_by, I)

  count_by <- data.frame(levels(sum_by[[var2]]), count_by)

  colnames(count_by) <- c(var2, "total")

  breakdown <- merge(sum_by, count_by, by = var2)
  breakdown$percentage <- breakdown$count / breakdown$total
  
  return(breakdown)
}

```

We see immediately that loans with 60-month terms have on average almost 4% 
higher interest than their shorter term counterparts. Continuing to plot versus
term, we find also that the longer terms are associated with significantly 
larger loan amounts as well. The median loan amount for 60 months is almost
double the amount taken in 36-month loans. Is the higher interest rate for 
longer loans a result of the loan simply being longer or because the loans are
larger (and thus riskier)?

```{r echo=FALSE, message=FALSE, warning=FALSE}

term_grade_breakdown <- percent_breakdown(working, "term", "grade")

ggplot(data = term_grade_breakdown, aes(x = grade, y = percentage)) +
  geom_bar(stat = "identity", aes(fill = term)) +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = c("0%", "25%", "50%", "75%", "100%")) +
  labs(y = "percentage of loans")

```

This plot makes it clear that the lower the grade, the higher the proportion of
loans that are of longer length. The difference is extremely evident with 
only 3.5% of grade A loans having a 60-month term but 87.5% of grade G loans
having the same term.

```{r echo=FALSE, message=FALSE, warning=FALSE}

grade_term_breakdown <- percent_breakdown(working, "grade", "term")

ggplot(data = grade_term_breakdown, aes(x = term, y = percentage)) +
  geom_bar(stat = "identity", aes(fill = grade)) +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = c("0%", "25%", "50%", "75%", "100%")) +
  labs(y = "percentage of loans")

```

Reversing the breakdown and instead plotting the percentage breakdown of each 
length of term by grade reveals similar information. 36-month loans are composed
mostly of grade A, B, and C loans (comprising a total 83%). Moving to 60-month
loans, the number of A and B grade loans drops off sharply (down 67%) while the
proportion of grade D, E, F, and G loans increases.

It is clear that, in general, Lending Club loans with 60-month terms are
classified as riskier with a lower grade. This is likely caused by both the
longer length of the loan and the much higher average value of the longer loans.

Is it really true though? Are 60-month loans more likely than their shorter 
counterparts to default or be charged off?

```{r echo=FALSE, message=FALSE, warning=FALSE}

term_status_breakdown <- percent_breakdown(
  subset(working, !(loan_status %in% c("Does not meet the credit policy. Status:Charged Off",
                              "Does not meet the credit policy. Status:Fully Paid"))),
  "loan_status", "term")

ggplot(data = term_status_breakdown, aes(x = term, y = percentage)) +
  geom_bar(stat = "identity", aes(fill = loan_status)) +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = c("0%", "25%", "50%", "75%", "100%")) +
  labs(y = "percentage of loans")

```

From this breakdown, we can see that the longer term loans don't necessary carry 
a higher risk for investors in terms of defaults and charge-offs. There are 
slightly more (around 1.5%) charge-offs for 60-month loans but not significantly.
While there are slightly less fully paid off loans for the longer term, this is
offset by a similarly higher number of currently in-progress and up to date loans.

It would seem that while Lending Club generally rates longer term loans as 
riskier with higher interest rates, from the current outcomes, it doesn't appear
as though 60-month loans are inherently more risky to investors in terms of
proportion of loans that default or are charged off.

***

In addition to the length of the term, let's examine what other factors go into
the determination of a loan's conditions.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = working, aes(x = annual_inc, y = int_rate)) +
  geom_point(alpha = 0.01, position="jitter") +
  coord_cartesian(
    xlim = c(quantile(working$annual_inc, probs = 0.01, na.rm = TRUE), 
             quantile(working$annual_inc, probs = 0.99, na.rm = TRUE)))

ggplot(data = working, aes(x = dti, y = int_rate)) +
  geom_point(alpha = 0.01, position="jitter") +
  coord_cartesian(
    xlim = c(quantile(working$dti, probs = 0.01, na.rm = TRUE), 
             quantile(working$dti, probs = 0.99, na.rm = TRUE)))

ggplot(data = working, aes(x = delinq_2yrs, y = int_rate)) +
  geom_point(alpha = 0.01, position="jitter") +
  coord_cartesian(
    xlim = c(quantile(working$delinq_2yrs, probs = 0.01, na.rm = TRUE), 
             quantile(working$delinq_2yrs, probs = 0.99, na.rm = TRUE)))

ggplot(data = working, aes(x = annual_inc, y = loan_amnt)) +
  geom_point(alpha = 0.01, position="jitter") +
  coord_cartesian(
    xlim = c(quantile(working$annual_inc, probs = 0.01, na.rm = TRUE), 
             quantile(working$annual_inc, probs = 0.99, na.rm = TRUE)))

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = working, aes(x = grade, y = int_rate)) +
  geom_boxplot()

ggplot(data = working, aes(x = grade, y = annual_inc)) +
  geom_boxplot() +
  coord_cartesian(
    ylim = c(quantile(working$annual_inc, probs = 0.01, na.rm = TRUE), 
             quantile(working$annual_inc, probs = 0.99, na.rm = TRUE)))

ggplot(data = working, aes(x = grade, y = dti)) +
  geom_boxplot() +
  coord_cartesian(
    ylim = c(quantile(working$dti, probs = 0.01, na.rm = TRUE), 
             quantile(working$dti, probs = 0.99, na.rm = TRUE)))

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

status_grade_breakdown <- percent_breakdown(
  subset(working, !(loan_status %in% c("Does not meet the credit policy. Status:Charged Off",
                              "Does not meet the credit policy. Status:Fully Paid"))), 
  "loan_status", "grade")

ggplot(data = status_grade_breakdown, aes(x = grade, y = count)) +
  geom_bar(stat = "identity", aes(fill = loan_status))

ggplot(data = status_grade_breakdown, aes(x = grade, y = percentage)) +
  geom_bar(stat = "identity", aes(fill = loan_status)) +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = c("0%", "25%", "50%", "75%", "100%")) +
  labs(y = "percentage of loans")

```

```{r echo=FALSE, message=FALSE, warning=FALSE}



```


# Multivariate Plots and Analysis

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!